Data Specialization for a Staged Lambda Calculus

### 2015-02-27 paper/splitting.tex
More edits on injections.

We now describe how to split \bbonem\ terms. The algorithm, given in
\cref{fig:termSplit}, proceeds recursively on the structure of~$e$. We
derive a few cases here.

\paragraph {Unit.} 
As a simple first case, consider splitting the unit value, $\tup{}$
which masks to $\mval {\tup{}} {\tup{}}$.
%Units trivial contain unit split into more units: $\tup{} \vsplito \mval {\tup{}} {\tup{}}$.
Splitting must produce a $c$ and $l.r$ satisfying,
\[
	\reduce c {\tup{\tup{},b}} \text{ and } [b/l]r \equiv \tup{}
\]
Our algorithm uses $c=\tup{\tup{},\tup{}}$ and $l.r = \_.\tup{}$.
Splitting $\pure{e}$ follows in similar fashion, except the immediate
result is instead the result of evaluating the \bbonep\ term $e$.

\paragraph {Injections.}
Now consider splitting the non-terminal, $\inl {e}$, where $\diaone
{e} \xi v$ and $\rtab \xi v \vsplito \mval i q$. Therefore:
\[
	\splitone e A c {l.r} \text{ and } \reduce c {\tup{i,b}} \text{ and } [b/l]r \equiv q
\]
Since $\diaone {\inl e} \xi {\inl v}$ and $\rtab \xi {\inl v} \vsplito \mval {\inl i} q$,
splitting must produce a $c'$ and $l.r'$ satisfying,
\[
	\reduce {c'} {\tup{\inl i,b'}} \text{ and } [b'/l]r' \equiv q
\]
Our algorithm simply emits the resumer of the subexpression ($l.r' =
l.r$), and generates a first-pass term $c'$ that produces the same
boundary value ($b$) as the subexpression. The entire first-pass term
is:
\[
c' = \letin {\tup {x,y}} c {\tup{\inl x, y}}
\]
\paragraph{Tuple.} For an example of splitting terms with more than one argument, consider the tuple $\tup{e_1,e_2}$,
where $\diaone {e_k} {\xi_k} {v_k}$ and $\rtab {\xi_k} {v_k} \vsplito
\mval {i_k} {q_k}$ for $k = 1,2$.  Inductively:
\[
	\splitone {e_k} A {c_k} {l_k.r_k} \text{ and } \reduce {c_k} {\tup{i_k,b_k}} \text{ and } [b_k/l_k]r_k \equiv q_k
\]
Since $\diaone {\tup{e_i,e_2}} {\xi_1,\xi_2} {\tup{v_1,v_2}}$ 
and $\rtab {\xi_1,\xi_2} {\tup{v_1,v_2}} \vsplito \mval {\tup{i_1,i_2}} {\tup{q_1,q_2}}$,
the validity of the split $\splitone {\tup{e_1,e_2}} A {c} {l.r}$ requires,
\[
	\reduce {c} {\tup{\tup{i_1,i_2},b}} \text{ and } [b/l]r \equiv \tup{q_1,q_2}
\]
To satisfy this, our algorithm uses a boundary value which is the tuple of the boundary values of the subexpressions,
$b = \tup{b_1,b_2}$. This is deconstructed with a tuple pattern, $l.r = \tup{l_1,l_2}.\tup{r_1,r_2}$,
and constructed using,
\[
c = \letin{\tup{y_1,z_1}}{c_1}{\letin{\tup{y_2,z_2}}{c_2}{\tup{\tup{y_1, y_2},\tup{z_1, z_2}}}}
\]

>>>>

Given the definition of masking, we now describe how to split
\bbonem\ terms. The algorithm, given in \cref{fig:termSplit}, proceeds
recursively on the structure of~$e$. We derive a few cases here.

\paragraph {Unit.} 
As a simple first case, consider splitting the unit value, $\tup{}$
which masks to $\mval {\tup{}} {\tup{}}$.
%Units trivial contain unit split into more units: $\tup{} \vsplito \mval {\tup{}} {\tup{}}$.
Splitting must produce a $c$ and $l.r$ satisfying,
\[
	\reduce c {\tup{\tup{},b}} \text{ and } [b/l]r \equiv \tup{}
\]
Our algorithm uses $c=\tup{\tup{},\tup{}}$ and $l.r = \_.\tup{}$.
Splitting $\pure{e}$ follows in similar fashion, except the immediate
result now becomes the result of evaluating the \bbonep\ term $e$.

\paragraph {Injections.}
Now consider splitting the non-terminal, $\inl {e}$, where $\diaone
{e} \xi v$ and $\rtab \xi v \vsplito \mval i q$. Therefore:
\[
	\splitone e A c {l.r} \text{ and } \reduce c {\tup{i,b}} \text{ and } [b/l]r \equiv q
\]
Since $\diaone {\inl e} \xi {\inl v}$ and $\rtab \xi {\inl v} \vsplito \mval {\inl i} q$,
splitting must produce a $c'$ and $l.r'$ satisfying,
\[
	\reduce {c'} {\tup{\inl i,b'}} \text{ and } [b'/l]r' \equiv q
\]
Our algorithm generates a first-pass term $c'$ that produces the same
boundary value ($b$) as the split of the subexpression $e$:
\[
c' = \letin {\tup {x,y}} c {\tup{\inl x, y}}
\]
Splitting also generates the same resumer as the split of the
subexpression ($l.r' = l.r$)

\paragraph{Tuple.} For an example of splitting terms with more than one argument, consider the tuple $\tup{e_1,e_2}$,
where $\diaone {e_k} {\xi_k} {v_k}$ and $\rtab {\xi_k} {v_k} \vsplito
\mval {i_k} {q_k}$ for $k = 1,2$.  Inductively:
\[
	\splitone {e_k} A {c_k} {l_k.r_k} \text{ and } \reduce {c_k} {\tup{i_k,b_k}} \text{ and } [b_k/l_k]r_k \equiv q_k
\]
Since $\diaone {\tup{e_i,e_2}} {\xi_1,\xi_2} {\tup{v_1,v_2}}$ 
and $\rtab {\xi_1,\xi_2} {\tup{v_1,v_2}} \vsplito \mval {\tup{i_1,i_2}} {\tup{q_1,q_2}}$,
the validity of the split $\splitone {\tup{e_1,e_2}} A {c} {l.r}$ requires,
\[
	\reduce {c} {\tup{\tup{i_1,i_2},b}} \text{ and } [b/l]r \equiv \tup{q_1,q_2}
\]
To satisfy this, our algorithm uses a boundary value which is the tuple of the boundary values of the subexpressions,
$b = \tup{b_1,b_2}$. This is deconstructed with a tuple pattern, $l.r = \tup{l_1,l_2}.\tup{r_1,r_2}$,
and constructed using,
\[
c = \letin{\tup{y_1,z_1}}{c_1}{\letin{\tup{y_2,z_2}}{c_2}{\tup{\tup{y_1, y_2},\tup{z_1, z_2}}}}
\]
